### 2009-03-30 text/paper.tex
Added paper stub.

\documentclass[12pt]{article}
\usepackage[utf8x]{inputenc}
\author{Magdalena Leshtanska \and Aleksandar Dimitrov}
\title{Error-Centric Annotation of Learner Corpora}
\begin{document}
\maketitle
\section{Introduction}\label{sec:intro}

Literature overview. Bla Bla.

All of the annotation schemes mentioned above focus on the \textbf{text} itself,
and will often even advise the annotator to \textit{modify} the underlying text
with the annotation. We believe that a more error-centric annotation of the data
can be beneficial.

Firstly, every annotation scheme that operates directly on the text or spans
over a piece of text will run into two kinds of problems:

\begin{itemize}
\item \textit{Interleaving annotations} occur when an error doesn't end before
another one begins. Given the tokens $a_1 a_2 a_3$, and two errors $e_1$ and
$e_2$ ranging over tokens $a_1$, $a_2$ and $a_2$ $a_3$ respectively, the
resulting markup will be confusing or outright impossible to read: $(e_1)a_1
(e_2)a_2(/e_1)a_3(/e_2)$. This is particularly a problem with XML,
since the specification\footnotemark explicitly disallows interleaving
markup.
\item \textit{Greedy annotation} covers tokens entirely uninvolved in the
\textit{``production''} of an error. If of the tokens $a_1 a_2 a_3$ only $a_1$
and $a_3$ are involved in the production of an error (say, an agreement error)
marking the entire token sequence as erroneous would involve blaming the
otherwise completely innocent token $a_2$. This can significantly increase the
noise in a corpus.
\end{itemize}
\footnotetext{Located at \texttt{http://www.w3.org/TR/REC-xml/}}

There is little data about the primary use cases for learner corpora known to
us, but it does not seem too far fetched to assume most linguists interested in
such data are not likely to browse through the corpora in order to find an
amusingly written piece of text. Instead, it seems to be safe to assume that
most would be interested in the \textbf{kinds of errors} that can occur in such
a corpus, and their particular properties. The error data should thus contain as
little noise as possible and be maximally specific while retaining a large amount
of generality.  It should furthermore be easily accessible, without having to
actually read the texts within the corpus itself.

Based on these assumptions, we decided to decouple the \textit{error markup}
from the \textit{corpus data}. Specifically, to our annotation scheme, the
errors and the text are two entirely different data structures. Every
particular error can reference tokens within the corpus using a \textbf{key},
similar to the way modern Relational Data Bases reference their data. This makes
the index more accessible and easier to maintain. Every marked up error contains exactly
the amount of information it needs in order to justify its presence, viz. by
defining what triggered its occurrence, and sufficient classification within the
taxonomy presented in \ref{sec:taxo}.

\section{Error Taxonomy and Annotation Scheme}\label{sec:scheme}

Learner errors are not easily classified due to their wide variety of
appearance, kind, and ambiguities involved. Typically, an annotation scheme will
strive to cover as large a scope it possibly can unambiguously classify, yet
there are many practical limitations to the potential coverage one can achieve.

One of the most prevalent restrictions is the need to devise the annotation
manual with great attention to detail, and to perform a meaningful assessment of
the quality of the annotated data after the annotation process. Even a most
meticulously well-defined manual is still going to be read and applied by
humans, who all have their personal intuition and character. It is easy to see
why this inherent individuality poses problems to the consistency of any kind of
annotated data. A well-written annotation manual can limit the effect this
natural variance can have on the data, but applying measures of consistency is
strictly necessary. The assessment of our data will be discussed in section
\ref{sec:results} while the remainder of this section will explain the rationale
behind our annotation manual, which is in turn given in section \ref{sec:manual}.

\subsection{Calling Errors by Name}\label{sec:taxo}

In order to identify errors occurring in learner language, one has to first
classify them. However, even then, identifying which class a particular error
belongs to is in no way an easy job. Devising an exhaustive taxonomy of errors
that can appear in natural language seems a daunting task, since science has so
far failed to regularize what \textit{is} a valid utterance of a language.
Therefore, our taxonomy will try to do some things only, and do them well. In
particular, we will not cover punctuation mistakes, since their possible
corrections can have cascading effects on the rest of the sentences and suddenly
give rise to all kinds of new errors.

Furthermore, our taxonomy will have a base case, called \textbf{gibberish} for
every error that cannot otherwise be classified, necessitated by its
deliberately established incompleteness. Sections delineated as such a case
should, however, be few in number, and as local as possible.

There are three other kinds of errors we want to term:

\begin{enumerate}
\item \textit{Spelling errors} are a primitive form of error. They contain no
meta-data or any kind of further specification except maybe a target word they
should have ended up as.
\item \textit{Form errors} are otherwise grammatical sentences which are,
however, semantically incorrect given the current context.
\item \textit{Grammar errors} simply do not belong to the target language because
of their erroneous morphology or syntax.
\end{enumerate}

\subsection{Representing Errors in Annotations}

\section{Annotation Manual}\label{sec:manual}

\subsection{Spelling Errors}\label{sec:speling}
A spelling error is an error that \textit{cannot} be explained by mistakes in
morphology, e.g. derivation or inflection. It will typically comprise only one
word. If it comprises more than one word, it is seen as a \textit{compound}
spelling error, i.e. two tokens separated by a space should have been
concatenated to yield only one token. Similarly, if the target hypothesis
accounts for more than one word, the token should have been split up into two.

\subsection{Form Errors}
Form errors occur in otherwise completely grammatical sentences, which, however
do not convey what was obviously intended by the author taking context into
account. Lexical errors are also form errors, if the given word is grammatically
valid within the sentence.

\subsection{Grammatical Errors}

\subsection{Devising Target Hypotheses}
Retain existing input by the author of the text to the greatest extent possible.
The text should be only minimally altered in the target hypothesis.

\section{Results, Assessment and Future Work}\label{sec:results}

\end{document}